\bf Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid

### 2015-02-18 Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid/main.tex
Pre my full review

In evolutionary optimization, and genetic algorithms~(GAs) in particular, search is often
biased through parameters. This can be beneficial as it allows practitioners to inject their
knowledge about the shape of the search landscape into the algorithm.
However, the quality of solutions found, and the speed at which they are found, is strongly tied to setting these parameters
correctly~\citep{goldberg:1991:gasize}. As such, either expert knowledge or exceedingly
expensive parameter tuning~\citep{grefenstette:1986:optimalga} are required to leverage
this feature to its fullest potential. Furthermore,  parameters such as population size, mutation rate, crossover
rate, tournament size, etc. usually have no clear relationship to the problem being solved, meaning even
domain experts may not understand how the parameters will interact with the problem or with each other.

@@

\subsection{Operation Counting}
\begin{figure}[t]
  \begin{centering}
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{rebuilds}
      \end{centering}
      \caption{Rebuilds}
      \label{fig-rebuilds}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{donations}
      \end{centering}
      \caption{Donations}
      \label{fig-donations}
    \end{subfigure}
  \end{centering}
  \caption{Computation costs incurred by model rebuilding (Figure~\ref{fig-rebuilds}) and
           repeated donations (Figure~\ref{fig-donations}) per evaluation as problem size increases.}
  \label{fig-costs}
\end{figure}

@@

Figure~\ref{fig-rebuilds} reports in an algorithmic sense how expensive model rebuilding is for search.
In order to calculate this value we recorded how many times search rebuilt the model during each run.
Under the minimal assumption that an evaluation costs $\Omega(N)$ time and that model rebuilds cost $O(N^2)$,
Figure~\ref{fig-rebuilds} shows the ratio cost of model rebuilding per evaluation. If the cost of model building
scaled linearly with evaluations, the relationship plotted for each problem should be asymptotically constant.
For Nearest Neighbor NK, Ising Spin Glasses, and Rastrigin this is the case. For both Trap problems
and HIFF there is slow growth in the ratio. The problem sizes used for MAX-SAT were not sufficient
to accurately gauge the asymptotic behavior. Together this suggests that while the cost of building the model
is almost linear per evaluation, it can grow slowly. However, even in the worst case (HIFF) this growth was
no more than twice the algorithmic cost of an evaluation even using 2048 bits.

@@

Figure~\ref{fig-small-pop} examines the effect of population size on LTGA's intermediate fitness by examining the effect
of reducing LTGA's population size to one tenth of the tuned value. The two problems shown are representative of the behavior of using
a smaller population size on the other five problems. Reducing the population size caused LTGA to improve earlier but plateau
at lower fitnesses. This caused LTGA's success rate to drop from 100 to 0 on Deceptive Step Trap and from 98 to 68 on Nearest
Neighbor NK.
Even when using a reduced population size for LTGA, P3 still achieved a fitness at
least as high as LTGA at 80\% of intervals.

>>>>

In evolutionary optimization, and genetic algorithms~(GAs) in particular, search is often
biased through parameters. This can be beneficial as it allows practitioners to inject their
knowledge about the shape of the search landscape into the algorithm.
However, the quality of solutions found, and the speed at which they are found, is strongly tied to setting these parameters
correctly~\citep{goldberg:1991:gasize}. As such, either expert knowledge or exceedingly
expensive parameter tuning~\citep{grefenstette:1986:optimalga} are required to leverage
this feature to its fullest potential. Furthermore,  parameters such as population size, mutation rate, crossover
rate, tournament size, etc. usually have no clear relationship to the problem being solved, meaning even
domain experts may not understand how the parameters will interact with the problem or with each other.
To further complicate matters, there is mounting evidence that parameter values should change
during search~\citep{goldman:2011:dynamic-parameters,laporte:2014:adaptivepop}.

@@

\subsection{Operation Counting}
\begin{figure}[t]
  \begin{centering}
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{rebuilds}
      \end{centering}
      \caption{Rebuilds}
      \label{fig-rebuilds}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{donations}
      \end{centering}
      \caption{Donations}
      \label{fig-donations}
    \end{subfigure}
  \end{centering}
  \caption{Estimated computation costs incurred by model rebuilding (Figure~\ref{fig-rebuilds}) and
           repeated donations (Figure~\ref{fig-donations}) per evaluation as problem size increases.}
  \label{fig-costs}
\end{figure}

@@

Figure~\ref{fig-rebuilds} reports in an algorithmic sense how expensive model rebuilding is for search.
In order to calculate this value we recorded how many times search rebuilt the model during each run.
Figure~\ref{fig-rebuilds} shows the estimated ratio of model rebuilding cost ($N^2$ per rebuild) over evaluation cost ($N$ per evaluation).
If the cost of model building
scaled linearly with evaluations, the relationship plotted for each problem should be asymptotically constant.
For Nearest Neighbor NK, Ising Spin Glasses, and Rastrigin this is the case. For both Trap problems
and HIFF there is slow growth in the ratio. The problem sizes used for MAX-SAT were not sufficient
to accurately gauge the asymptotic behavior. Together this suggests that while the cost of building the model
is almost linear per evaluation, it can grow slowly. However, even in the worst case (HIFF) this growth was
no more than twice the algorithmic cost of an evaluation even using 2048 bits.

@@

Figure~\ref{fig-small-pop} examines the effect of population size on LTGA's intermediate fitness by
reducing LTGA's population size to one tenth of the tuned value. The two problems shown are representative of the behavior of using
a smaller population size on the other five problems. Reducing the population size caused LTGA to improve earlier but plateau
at lower fitnesses. This caused LTGA's success rate to drop from 100 to 0 on Deceptive Step Trap and from 98 to 68 on Nearest
Neighbor NK.
Even when using a reduced population size for LTGA, P3 still achieved a fitness at
least as high as LTGA at 80\% of intervals.

### 2015-02-18 Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid/reviews.tex
Pre my full review

We agree, though our overall point is more general in that we are trying to get the community to rethink \textbf{all} parameter settings. In the original submission Section 9, Section 10.2,  Figure 11, and Figure 13a explicitly discuss the benefits we believe P3 achieves due to side stepping the fixed size generational scheme. We also had previously mentioned it in the third paragraph of the conclusion. Nonetheless, we have now added some additional references and discussion expanding those sections
%TODO Add references and discussion

@@

It is a busy figure. We have added more about the figure and its conclusions to the text.
%TODO add explanation

@@

However, adding initial local search as a tunable parameter to hboa and LTGA when partial evaluations are not available is not part
of existing work for either method. Doing so would represent novel research outside the scope of our topic, and provide those techniques
with even more problem specific configuration. We welcome future work by other researchers exploring the possible effects of combining
local search with their techniques.

>>>>

We agree, though our overall point is more general in that we are trying to get the community to rethink
\textbf{all} parameter settings. In the original submission Section 9, Section 10.2,
Figure 11, and Figure 13a explicitly discuss the benefits we believe P3 achieves due
to side stepping the fixed size generational scheme.
We also had previously mentioned it in the third paragraph of the conclusion.
Nonetheless, we have now added some additional references and discussion to the introduction to indicate its importance.

@@

It is a busy figure. We have worked to clarify the text describing the figure.

@@

However, adding initial local search as a tunable parameter to hboa and LTGA when partial evaluations are not available is not part
of existing work for either method. Doing so would also provide those techniques
with even more problem specific configuration. We welcome future work by other researchers exploring the possible effects of combining
local search with their techniques.
